Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2019-01-05 review-responses.tex

\begin{point}{}
    The authors identify two sources of noise: Sampling noise (arising from
    sampling of individuals) and Process noise (stochasticity of population
    history).  Using SLiM, which I gather controls for random variation in the
    realized pedigree and hence eliminates (or only reduces?) Process noise,
    the authors obtain estimates for the relative magnitude of the two sources
    of noise.  This is useful but seems incomplete for two reasons:
    \begin{enumerate}
        \item       A third source of noise is sampling a finite number of
            genes or SNPs.  To some extent one can trade off sampling
            individuals vs loci, but there are limits to how effective that is
        \item       The relative magnitudes of the three sources of error are
            not constant and can vary dramatically depending on parameter
            values.  Process error is inversely related to Ne but otherwise
            fixed in magnitude; it does not change when sampling errors change.
            This means that Process error becomes relatively more important
            when there is a lot of replication in sampling.  In particular,
            genetic sampling error gets very small as the number of SNPs
            becomes arbitrarily large, in which case it is dwarfed by Process
            error (e.g., see Figure 6 in our 2009 Mol Ecol paper cited by the
            authors).
    \end{enumerate}
\end{point}

\reply{
}

\begin{point}{}
    Also in relation to genomics-scale datasets, two general potential
    advantages can be identified:
    \begin{enumerate}
        \item       Arbitrarily large numbers of loci can increase precision
            and statistical power for existing analyses;
        \item       Ability to access information related to genome structure
            and function can provide novel insights and make possible
            qualitatively new types of analyses
    \end{enumerate}
    Lots of people have genomics-scale datasets these days so some discussion of
    this issue seems important.
\end{point}

\reply{
}

\begin{point}{}
    ``Can our method replace resistance-based methods? Perhaps, but the substantial
    uncertainty we saw on graphs with only tens of nodes is indicative of a
    larger problem we face for realistic models. We have seen that
    discretization of space results in substantial modeling error, in part
    because of randomness of geographic sampling and unmodeled process noise
    can lead to overfitting. Partitioning space into a finer grid should help
    with these problems, but tends to make the inference problem itself more
    ill-conditioned: with more connections, changing the value of one
    connection affects coalescence time less, and so inferences about that
    value must necessarily be less certain. This tradeoff implies some degree
    of unavoidable uncertainty.  Reproducible, reliable inference will likely
    require development of new inference methods that explicitly model
    continuous geography. `` This seems like a fairly crucial paragraph.  But
    what exactly is the take-home message for biologists interested in
    practical application of the methodology?
\end{point}

\reply{
}

\begin{point}{}
    Coalescent rates are critical to model performance, which can be poor if the
    rate is too low.  It is obvious that the coalescent rate is inversely
    related to Ne, but in practical terms, what does it mean when the
    coalescent rate is lowered from 10 to 0.1?
\end{point}

\reply{
}

\begin{point}{\revref}
    ``we varied the standard deviation of the noise from 1/1000th to 1/50th of
    the mean value of C for that grid.''  This sounds like a CV of 0.001 to
    0.02 for noise, both of which would be exceedingly small.  If that is not
    what is meant, please clarify.
\end{point}

\reply{
}

\begin{point}{\revref}
    ``if fecundity is variable'' -- on what scale?  Among populations?  Among
    individuals?  Among ages within the same individual? 
\end{point}

\reply{
}

\begin{point}{\revref}
    ``It is not required in this approach to have samples from every spatial
    location ...''  Required for what?  Are they saying that unsampled but real
    (ghost) populations have no effect on precision or bias?
\end{point}

\reply{
}

\begin{point}{}
    Many of the figures are too small to read easily without a magnifying glass.
    Figures 3-5 are large enough but they include miniscule fonts that make it
    impossible to tell exactly what is plotted.  Green and black do not provide
    enough contrast.
\end{point}

\reply{
}

\begin{point}{\revref}
    another limitation of direct observations is that they only provide
    information about movement of individuals but not gene flow, which requires
    successful breeding
\end{point}

\reply{
}

\begin{point}{\revref}
    the ``wrong answer'' compared to what?
\end{point}

\reply{
}

\begin{point}{\revref}
    What is EEMS?
\end{point}

\reply{
}

\begin{point}{\revref}
    What exactly is a more general problem?
\end{point}

\reply{
}

\begin{point}{\revref}
    need a semicolon, not a comma, after ``two genomes''
\end{point}

\reply{
    We reworded this.
}

\begin{point}{184:}
    why quotes for ``from the model''?
\end{point}

\reply{
}

\begin{point}{}
    ``suppose that the spatial arrangement of populations is given by a graph
    ...''  ``Graph'' is a very general term and not clear exactly what is meant
    here.  If the word has a more specific meaning in this context, please
    define it.
\end{point}

\reply{
}

\begin{point}{Fig 6:}
    the barrier is ``sufficiently thick'' to prohibit migration.  ``Thick''
    implies distance, but is that a requirement?  Couldn’t the barrier be
    higher or deeper or less penetrable for a different reason than thickness?
\end{point}

\reply{
}

\reviewersection{4}

\begin{quote}
    This is a polished theoretical and methodological contribution towards
    understanding the genetic processes that affect genetic distance and thus
    connectivity of species across the landscape. The authors make a
    little-appreciated point about considering the coalescent time as a
    predictor of genetic distance as opposed to resistance distance which can
    result in spurious results  where genetic distance is shaped by
    asymmetrical gene flow and coalescent processes.

    There are some points of clarification however that I think need to be made
    so that the paper is accessible and clear to its intended audience, and to
    ensure the paper can be adopted and influence analyses in this area. The
    differences in the highly spatial resistance distance approach versus the
    genetics-based coalescent time-based approach need to be distinguished more
    clearly, as in some ways this may be seen as comparing apples and oranges.
    Issues around timescales of genetic change (effects of landscape vs
    coalescent times) need to be further addressed and elucidated.  The
    differences in the results gained from the approaches are clearly defined
    and presented however.

    The types of questions and the type of demographic scenarios under which
    resistance vs. coalescense based approaches may be an improvement, or
    combined in the same study should be made more clearly. I suggest a
    summarising table or flow chart to this effect, that addresses time scales
    also.
\end{quote}

\begin{point}{Line 7 of abstract and generally:}
This statement is conditional on the characteristics of the data you have, and
    your study aim. If the study aim is to  ask how landscape features affect
    genetic distance, then the coalescent time is not going to explicitly test
    this, or? It needs to be made clear how different study objectives will
    benefit in landscape genetics and under what circumstances a 'more
    accurate' predictor of genetic distance can be gained. It seems appropriate
    to mention that this approach would greatly complement existing landscape
    genetic analysis approaches to see the relative effects of resistance and
    coalescent time. Implications for management decisions based on resistance
    models could be mentioned, as resistance models are intended to inform
    conservation decisions in some cases.
\end{point}

\reply{
}

\begin{point}{Line 26-38:}
    Extremely long sentence, please divide.
\end{point}

\reply{
    Not sure what they mean.
}

\begin{point}{\revref}
    Would it make sense to select genetic markers that are more informative for
    coalescence rates by narrowing the level of variation accepted in the
    model? Would this help error rates justifiably or just bias interpretation?
\end{point}

\reply{
}

\begin{point}{\revref}
    I think it  might be helpful if this point was made earlier and reiterated
\end{point}

\reply{
}

\reviewersection{5}

\begin{quote}
    This article focuses on the inference of gene flows between populations from
    the analysis of georeferenced genetic data. In particular, a new
    coalescent-based technique, which applies to the analysis of pairs of
    sampled individuals, is compared to an approach based on "resistance
    distances" which aim at approximating coalescent distances. This work is
    interesting because resistance distances are at the core of several popular
    techniques in landscape genetics. It is thus important to gain a good
    understanding of how good this phenomenological approximation is compared
    to more mechanistic techniques such as the structured coalescent. The
    article clarifies the connections between the two approaches in a
    convincing manner. It thereby fully deserves to be published in Molecular
    Ecology Resources. There are however a few points (see below) that I think
    should be improved should the authors see fit.
\end{quote}

% Section "Hitting times of Markov chains"

\begin{point}{}
    $Pr(X_{t+e}=y | X_t=y)$ should read $Pr(X_{t+e}=y | X_t=x)$.
\end{point}

\reply{
}

\begin{point}{}
    Regarding Equation 2, I would recommend giving more detail explaining where
    this comes from. Here is a suggestion:
    \begin{align*}
        H_xz &= dt + \sum_y Pr(X_{dt} = y | X_0 = x) H_{yz} \\
             &= dt + \sum_{y != x} Q_{xy} dt H_{yz} + (Q_{xx} dt + 1) H_{xz}
    \end{align*}
\end{point}

\reply{
}

\begin{point}{}
    On a related note, it might be helpful to give more explanations about Equation 6. In
    fact, I am not quite sure I agree with it. Indeed, for $x \neq z$, we have
    \begin{align*}
        C_xz &= dt + {} \\
                & \qquad
             1/2 \sum_y Pr(X_{dt} = y | X_0 = x) C_{yz} + {} \\
                & \qquad
             1/2 \sum_y Pr(X_{dt} = y | X_0 = z) C_{xy}, \\
           &= dt + {} \\
                & \qquad
             1/2 \sum_{y != x} Q_{xy} dt C_{yz} + (Q_{xx} dt + 1) C_{xz} + \\
                & \qquad
             1/2 \sum_{y != x} Q_{zy} dt C_{xy} + (Q_{xx} dt + 1) C_{xz},
    \end{align*}
    where the 1/2 terms correspond to the probability of "selecting" the lineage found in x
    (resp. z) at time 0 to move to y after dt. The last equation above then gives
    \begin{align*}
        -2 = \sum_y Q_{xy} C_{yz} + Q_{zy} C_{xy}
    \end{align*}
    which differs from the one given in the ms. I have probably missed something in the derivation
    used by the authors. Giving a few more explanations here might help avoiding any confusion.
\end{point}

\reply{
}

% Section "Bayesian inference of movement rates"

\begin{point}{}
    I was also a bit puzzled by the likelihood function that is at the core of the
    Bayesian inference proposed here (see Equation 8). The spatial component of
    the data seems completely hidden. It is in fact conveyed by the coalescent
    times $C_{ij}(G,\gamma)$ but a more detailed description of the likelihood
    function where both genetic and spatial information appear explicitly would
    facilitate the understanding of the proposed inference technique.
\end{point}

\begin{point}{}
    Moreover, my understanding is that the joint probability of the pairwise genetic distances
    ($D$) and spatial locations ($L$) given $G$ and $\gamma$, i.e.,
    $p(D,L|G,\gamma)$, is expressed as the product 
    $p(D|L,G,\gamma) \times p(L|G,\gamma)$.  Thanks to Equation (7) the first density is in fact
    $p(D|C)$, where $C$ are the coalescent times. The authors use a normal distribution for $D|C$ but
    nothing is said  about the distribution of $L|G,\gamma$. It is paramount to explicitly define
    every probabilistic distribution in the likelihood of the model used for the inference so as
    to clarify what are the underlying simplifying hypotheses the proposed approach relies on.
\end{point}

\begin{point}{}
    On a related note, my understanding is that the uncertainty affecting
    coalescent times conditional on the movement rates is not taken into
    account in the proposed approach (i.e., knowing $G$, one gets $C$ through
    Equation 7 in a deterministic fashion). I would recommend that the authors
    discuss this point.
\end{point}

\reply{
}

% Section "Identifying a barrier to gene flow"

\begin{point}{}
    The sentence "with noise standard deviation equal to 1/1000th of the mean
    coalescence time (left) and 1/100th of the mean coalescence time (right)"
    should read "with noise standard deviation equal to 1/1000th of the mean
    coalescence time (top) and 1/100th of the mean coalescence time (bottom)"
\end{point}

\reply{
}

% Section "Discussion"

\begin{point}{}
    The authors call for new models to be developed that would properly accommodate
    for individuals distributed in space along a continuum as opposed to
    discrete demes, as is the case with the structured coalescent. The spatial
    Lambda-Fleming Viot model is one of these. It has been studied extensively
    from a mathematical standpoint (see the body of work by Etheridge, Barton
    and Veber amongst others) and has been used for the inference using a
    Bayesian approach (see Guindon, Guo, Welch, Theoretical Population Biology,
    2016; but also Koskela's PhD thesis
    \url{http://wrap.warwick.ac.uk/88065/1/WRAP_Theses_Koskela_2016.pdf}).  The
    spatial Lambda-Fleming-Viot model is mechanistic and enjoys many good
    statistical and biological properties (e.g., it does not suffer from the
    ``clumping'' issue that affects the Wright-Malecot model).
\end{point}

\reply{
}

\begin{point}{}
    Assuming that space is discrete in arguably the biggest weakness of the structured
    coalescent. It is quite often not biologically realistic and breaks down
    mathematically when considering larger numbers of smaller demes. The
    spatial Lambda-Fleming-Viot avoids this important shortcoming and thus
    probably deserves more attention.
\end{point}

\reply{
}
